Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-07-07 paper/paper.tex
replace A^2M to AAM

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Abstracting abstract machines is a systematic methodology for constructing
  sound static analyses for higher-order languages, by deriving small-step \emph{abstract}
  abstract machines (A\textsuperscript{2}Ms) that perform abstract interpretation from abstract machines that
  perform concrete evaluation. Darais et al. apply the same underlying idea to monadic definitional interpreters,
  and obtain monadic \emph{abstract} definitional interpreters (ADIs) that perform
  abstract interpretation in big-step style using monads.
  Yet, the relation between small-step abstract abstract machines and big-step
  abstract definitional interpreters is not well studied.

  In this paper, we explain their functional correspondence and demonstrate how to
  syntactically transform small-step abstract abstract machines into big-step abstract
  definitional interpreters.
  Building on known semantic interderivation techniques from the concrete evaluation
  setting, the transformations include linearization, lightweight fusion, disentanglement,
  refunctionalization, and the left inverse of the CPS transform.
  Linearization expresses nondeterministic choice through first-order data types,
  after which refunctionalization transforms the first-order data types that represent continuations
  into higher-order functions.
  The refunctionalized A\textsuperscript{2}M is an abstract interpreter written in continuation-passing style (CPS)
  with two layers of continuations, which can be converted back to direct style with delimited
  control operators. Based on the known correspondence between delimited control and monads,
  we demonstrate that the explicit use of monads in abstract definitional interpreters
  is optional.

@@

In addition, those abstract abstract machines with an unbounded stack naturally correspond
to abstract definitional interpreters. We show that refunctionalizing an A\textsuperscript{2}M with an
unbounded stack (and utilizing the proper caching algorithm) leads to a pushdown control-flow analysis.

@@

       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & & & &
       \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
     };
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-4);
     \chainin (m-2-5) [join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-5);
     \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{CPS trans.}}}];
   }
   {
     [start chain, color=gray] \chainin (m-1-5);
     \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{inline monads}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-1);
     \chainin (m-2-2) [join={node[above,labeled] {
       \begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{lightw. fusion} \end{smallmatrix}
     }}];
     \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-2);
     \chainin (m-2-1) [join={node[below,labeled] {\begin{smallmatrix} \text{splitting} \\ \text{+} \\
     \text{replacing cont.} \\ \text{by list} \end{smallmatrix}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-3);
     \chainin (m-2-2) [join={node[below,labeled] {\text{entang.}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-3);
     \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-4);
     \chainin (m-2-3) [join={node[below,labeled] {\text{defunc.}}}];
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-1);
     { [start branch=A] \chainin (m-2-1) [join={node[left, labeled] {\alpha}}]; }
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-5);
     {
       [start branch=B] \chainin (m-2-5)
       [join={node[right, labeled] {\alpha}}];
     }
     {
       [start branch=D, color=gray] \chainin (m-3-1)
       [join={node[right, labeled] {\text{}}}];
     }
   }
   {
     [start chain, color=gray] \chainin (m-3-1);
     {
       [start branch=C] \chainin (m-3-5)
       [join={node[above, labeled] {\text{Functional correspondences for concrete semantic artifacts}}}];
     }
   }
   \draw[gray,dashed,transform canvas={xshift=0.5ex}] (m-1-5.north west) -- (m-2-5.south west);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-1-5.north west) -- (m-1-5.north east);
   \draw[gray,dashed,transform canvas={xshift=-0.2ex}] (m-1-5.north east) -- (m-2-5.south east);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-2-5.south west) -- (m-2-5.south east);
 \end{tikzpicture}
 \caption{Functional correspondence between and transformations from A\textsuperscript{2}Ms to ADIs} \label{fig:trans}
\end{center}
\end{figure}

@@

We begin by reviewing necessary background in Section~\ref{background},
as well as by introducing some of the basic code structures used throughout the paper.
We then address the main contribution of this paper: bridging the gap between
small-step A\textsuperscript{2}Ms and big-step ADIs by applying a series of well-known systematic transformations
drawn from concrete semantic artifacts.
Those transformations are shown in Figure~\ref{fig:trans} and summarized here, with full
descriptions in their respective associated sections:

@@

  \item \textbf{Lightweight fusion}~\cite{DANVY2008100,Ohori:2007:LFF:1190216.1190241}:
    We apply a fusion transformation to the linearized A\textsuperscript{2}M, which
    combines the single-step function @step@ and the
    driving function into one, but keeps all the machine state representations intact
    (Section~\ref{fusing}).

@@

  \item \textbf{Refunctionalization}~\cite{DANVY2009534,Danvy:2006:RW:2171265.2171268}:
    We apply refunctionalization to the disentangled A\textsuperscript{2}M, which
    %as the left inverse of Reynolds' defunctionalization \cite{Reynold:72,Reynolds:HOSC98-revisited}
    replaces the first-order data types representing continuations with
    higher-order functions, and associates dispatching logic with proper higher-order functions.
    We obtain an abstract interpreter written in continuation-passing style with an
    additional layer of continuations due to nondeterminism.

    For clarity, we first present a vanilla version of a refunctionalized A\textsuperscript{2}M which simply
    converts the stack structures to higher-order functions but keeps other parts unchanged.
    We then adopt a caching algorithm \cite{darais2017abstracting} to guarantee the termination
    of abstract interpretation.
    At the end of this section, we review the pushdown control-flow analysis and examine how
    computable and precise call/return matching is obtained through these transformations
    (Section~\ref{sec:refunc}).

  \item \textbf{Direct-style transformation}~\cite{DBLP:journals/scp/Danvy94}:
    We finally transform the refunctionalized A\textsuperscript{2}M back to a
    direct-style interpreter by using delimited control operators
    (Section~\ref{directstyle}).
\end{itemize}

@@

Similar to the CESK machine, the machine state of an A\textsuperscript{2}M has a control, an environment, a store,
and continuation, as well as a timestamp. However, there are several notable differences
between an A\textsuperscript{2}M's store and the CESK machine's store. In AAM, the store maps addresses to sets of
values; it stores all possible values for a particular address. As such, dereferencing
addresses becomes nondeterministic. Also, the store performs \emph{joining}, rather than
overwriting, when updating elements. Furthermore, the continuations are likewise allocated
on the store instead of formed into a linked list, and the continuation component becomes
an address that maps to a set of continuations in the store instead of directly embedded
in the state.

@@

The pushdown AAM with an unbounded stack is also described by \citeauthor{van2012systematic}
\cite{van2012systematic}.
We show it here is to maintain consistency of abstract semantics during
transformations -- since the eventual abstract definitional interpreters simply inherit
a precise control-flow from their defining language, we also would like to start from a
pushdown A\textsuperscript{2}M that is also able to precisely match calls and returns.
For readers who are not familiar with pushdown analysis, we provide a detailed discussion
in Section~\ref{pdcfarevisit}.

@@

In the previous section, we show that by keeping an unabstracted stack in the state
space, we can recover the precise call/return match. Now we begin describing the
transformations step by step. Our base machine is the A\textsuperscript{2}M with an unbounded
stack; as we will later transform the stack structure to higher-order functions
representing continuations, it does not matter what kind of A\textsuperscript{2}M we start from.
This is because it would not be possible to construct a new frame on the stack, nor
to allocate continuations in the store after refunctionalization.
Thus, our choice to start from an unbounded-stack A\textsuperscript{2}M is motivated simply because
it has an equivalent stack model to abstract definitional interpreters (our final target).

@@

With fusing complete, our A\textsuperscript{2}M appears similar to a ``big-step'' interpreter, although it
still has the machine state representation inside. The inverse transformation of this step
would be splitting the work-list algorithm apart to an individual function.

@@

The purpose of disentanglement is to reveal the defunctionalized form of our A\textsuperscript{2}M.
Functions @continue@ and @ndcontinue@ that dispatch different data types representing
continuations are disentangled from @drive_step@.
The inverse transformation of this step would be simply inlining @ndcontinue@ into @continue@,
and then inlining @continue@ into @drive_step@.

@@

The continuation and meta-continuation are added as arguments to @aeval@.
Before jumping into the details of refunctionalizing the stack structure and constructing
@new_cont@ and @new_mcont@, we have two important observations in the disentangled A\textsuperscript{2}M
where the continuations were represented by data types:
1) When constructing the @new_ndk@, we observe that it explicitly uses @new_frames@ as one
argument of its top @NDCont@ object. The higher-order version of these continuations follows
the same pattern.
2) In the meta-continuation dispatching function @ndcontinue@, the size of the closures set is
decreasing every time while the other parts of meta-continuation are kept. This means in
the higher-order version, we also need an ability to update the top frame of meta-continuation
and keep the rest as the same.

@@

In this section, based on the refunctionalized A\textsuperscript{2}M, we
further present several other transformations toward a big-step abstract
definitional interpreter:

@@

\textit{Call/Return Matching through Refunctionalization and Caching.}
The refunctionalized A\textsuperscript{2}M we obtained has the perfect match on return-flows even though
we do not have a stack model. The higher-order functions representing continuations
already connect all the executions in order. In fact, the continuations (call stack)
of the analyzed language are blended into the call stack of our defining language
(Scala) through refunctionalization, and hence the call/return flow of the analyzed
language is naturally matched.

In fact, we started with an unbounded-stack A\textsuperscript{2}M which already precisely
matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
semantic artifact; the consequence is that we have no place to explicitly save the context
information in the store or on the stack.
However, refunctionalization not only forces the nondeterministic control flow to be higher-order,
but also drives us to use a different caching algorithm.

@@

%probably Store Widening \& Abstract Garbage Collection
\section{Back to Direct-Style} \label{directstyle}

In the previous section, we presented a refunctionalized A\textsuperscript{2}M in extended continuation-passing style,
and then simplified it by utilizing the @nd@ operator.
To obtain a definitional abstract interpreter in direct-style, we have multiple choices of how to proceed:

@@

\Citet{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
for small-step A\textsuperscript{2}M to achieve pushdown analysis. This method is both
simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
Based on the A\textsuperscript{2}M we presented in Section \ref{aam}, the only changes in the code required is not
only keeping track of the entry-point expression of the callee, but also holding the target
environment when allocating a continuation addresses. No other pieces of code would need to be
modified. Notably, this change only causes a constant-factor increase in time complexity of
the analysis if the store is widened.

@@

This paper is greatly inspired by the work of~\citet{darais2017abstracting}.
A first surface-level difference is that our presentation uses A-Normal Form $\lambda$-calculus,
instead of plain $\lambda$-calculus, although our work could be easily adapted to handle plain
$\lambda$-calculus as well. Going through the transformation steps from A\textsuperscript{2}Ms to ADIs instead of
only considering the final ADI result provides additional insights. In particular, it reveals that
refunctionalization plays an important role for inheriting the call structure (i.e., the model of
the stack) from the meta-language.
In addition, we demonstrate that the explicit use of monads in ADIs is optional, by translating to
direct-style ADIs with delimited control operators in addition to the monadic version of \citet{darais2017abstracting}.
We use the delimited control operators @shift@ and @reset@ \cite{Danvy:1990:AC:91556.91622} as implemented in Scala
\cite{rompf2009implementing}. The correspondence of delimited continuations and monads is well known
\cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions, filinski1994representing}.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Abstracting abstract machines is a systematic methodology for constructing
  sound static analyses for higher-order languages, by deriving small-step \emph{abstract}
  abstract machines (AAMs) that perform abstract interpretation from abstract machines that
  perform concrete evaluation. Darais et al. apply the same underlying idea to monadic definitional interpreters,
  and obtain monadic \emph{abstract} definitional interpreters (ADIs) that perform
  abstract interpretation in big-step style using monads.
  Yet, the relation between small-step abstract abstract machines and big-step
  abstract definitional interpreters is not well studied.

  In this paper, we explain their functional correspondence and demonstrate how to
  syntactically transform small-step abstract abstract machines into big-step abstract
  definitional interpreters.
  Building on known semantic interderivation techniques from the concrete evaluation
  setting, the transformations include linearization, lightweight fusion, disentanglement,
  refunctionalization, and the left inverse of the CPS transform.
  Linearization expresses nondeterministic choice through first-order data types,
  after which refunctionalization transforms the first-order data types that represent continuations
  into higher-order functions.
  The refunctionalized AAM is an abstract interpreter written in continuation-passing style (CPS)
  with two layers of continuations, which can be converted back to direct style with delimited
  control operators. Based on the known correspondence between delimited control and monads,
  we demonstrate that the explicit use of monads in abstract definitional interpreters
  is optional.

@@

In addition, those abstract abstract machines with an unbounded stack naturally correspond
to abstract definitional interpreters. We show that refunctionalizing an AAM with an
unbounded stack (and utilizing the proper caching algorithm) leads to a pushdown control-flow analysis.

@@

       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & & & &
       \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
     };
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-4);
     \chainin (m-2-5) [join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-5);
     \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{CPS trans.}}}];
   }
   {
     [start chain, color=gray] \chainin (m-1-5);
     \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{inline monads}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-1);
     \chainin (m-2-2) [join={node[above,labeled] {
       \begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{lightw. fusion} \end{smallmatrix}
     }}];
     \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-2);
     \chainin (m-2-1) [join={node[below,labeled] {\begin{smallmatrix} \text{splitting} \\ \text{+} \\
     \text{replacing cont.} \\ \text{by list} \end{smallmatrix}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-3);
     \chainin (m-2-2) [join={node[below,labeled] {\text{entang.}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-3);
     \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-4);
     \chainin (m-2-3) [join={node[below,labeled] {\text{defunc.}}}];
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-1);
     { [start branch=A] \chainin (m-2-1) [join={node[left, labeled] {\alpha}}]; }
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-5);
     {
       [start branch=B] \chainin (m-2-5)
       [join={node[right, labeled] {\alpha}}];
     }
     {
       [start branch=D, color=gray] \chainin (m-3-1)
       [join={node[right, labeled] {\text{}}}];
     }
   }
   {
     [start chain, color=gray] \chainin (m-3-1);
     {
       [start branch=C] \chainin (m-3-5)
       [join={node[above, labeled] {\text{Functional correspondences for concrete semantic artifacts}}}];
     }
   }
   \draw[gray,dashed,transform canvas={xshift=0.5ex}] (m-1-5.north west) -- (m-2-5.south west);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-1-5.north west) -- (m-1-5.north east);
   \draw[gray,dashed,transform canvas={xshift=-0.2ex}] (m-1-5.north east) -- (m-2-5.south east);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-2-5.south west) -- (m-2-5.south east);
 \end{tikzpicture}
 \caption{Functional correspondence between and transformations from AAMs to ADIs} \label{fig:trans}
\end{center}
\end{figure}

@@

We begin by reviewing necessary background in Section~\ref{background},
as well as by introducing some of the basic code structures used throughout the paper.
We then address the main contribution of this paper: bridging the gap between
small-step AAMs and big-step ADIs by applying a series of well-known systematic transformations
drawn from concrete semantic artifacts.
Those transformations are shown in Figure~\ref{fig:trans} and summarized here, with full
descriptions in their respective associated sections:

@@

  \item \textbf{Lightweight fusion}~\cite{DANVY2008100,Ohori:2007:LFF:1190216.1190241}:
    We apply a fusion transformation to the linearized AAM, which
    combines the single-step function @step@ and the
    driving function into one, but keeps all the machine state representations intact
    (Section~\ref{fusing}).

@@

  \item \textbf{Refunctionalization}~\cite{DANVY2009534,Danvy:2006:RW:2171265.2171268}:
    We apply refunctionalization to the disentangled AAM, which
    %as the left inverse of Reynolds' defunctionalization \cite{Reynold:72,Reynolds:HOSC98-revisited}
    replaces the first-order data types representing continuations with
    higher-order functions, and associates dispatching logic with proper higher-order functions.
    We obtain an abstract interpreter written in continuation-passing style with an
    additional layer of continuations due to nondeterminism.

    For clarity, we first present a vanilla version of a refunctionalized AAM which simply
    converts the stack structures to higher-order functions but keeps other parts unchanged.
    We then adopt a caching algorithm \cite{darais2017abstracting} to guarantee the termination
    of abstract interpretation.
    At the end of this section, we review the pushdown control-flow analysis and examine how
    computable and precise call/return matching is obtained through these transformations
    (Section~\ref{sec:refunc}).

  \item \textbf{Direct-style transformation}~\cite{DBLP:journals/scp/Danvy94}:
    We finally transform the refunctionalized AAM back to a
    direct-style interpreter by using delimited control operators
    (Section~\ref{directstyle}).
\end{itemize}

@@

Similar to the CESK machine, the machine state of an AAM has a control, an environment, a store,
and continuation, as well as a timestamp. However, there are several notable differences
between an AAM's store and the CESK machine's store. In AAM, the store maps addresses to sets of
values; it stores all possible values for a particular address. As such, dereferencing
addresses becomes nondeterministic. Also, the store performs \emph{joining}, rather than
overwriting, when updating elements. Furthermore, the continuations are likewise allocated
on the store instead of formed into a linked list, and the continuation component becomes
an address that maps to a set of continuations in the store instead of directly embedded
in the state.

@@

The pushdown AAM with an unbounded stack is also described by \citeauthor{van2012systematic}
\cite{van2012systematic}.
We show it here is to maintain consistency of abstract semantics during
transformations -- since the eventual abstract definitional interpreters simply inherit
a precise control-flow from their defining language, we also would like to start from a
pushdown AAM that is also able to precisely match calls and returns.
For readers who are not familiar with pushdown analysis, we provide a detailed discussion
in Section~\ref{pdcfarevisit}.

@@

In the previous section, we show that by keeping an unabstracted stack in the state
space, we can recover the precise call/return match. Now we begin describing the
transformations step by step. Our base machine is the AAM with an unbounded
stack; as we will later transform the stack structure to higher-order functions
representing continuations, it does not matter what kind of AAM we start from.
This is because it would not be possible to construct a new frame on the stack, nor
to allocate continuations in the store after refunctionalization.
Thus, our choice to start from an unbounded-stack AAM is motivated simply because
it has an equivalent stack model to abstract definitional interpreters (our final target).

@@

With fusing complete, our AAM appears similar to a ``big-step'' interpreter, although it
still has the machine state representation inside. The inverse transformation of this step
would be splitting the work-list algorithm apart to an individual function.

@@

The purpose of disentanglement is to reveal the defunctionalized form of our AAM.
Functions @continue@ and @ndcontinue@ that dispatch different data types representing
continuations are disentangled from @drive_step@.
The inverse transformation of this step would be simply inlining @ndcontinue@ into @continue@,
and then inlining @continue@ into @drive_step@.

@@

The continuation and meta-continuation are added as arguments to @aeval@.
Before jumping into the details of refunctionalizing the stack structure and constructing
@new_cont@ and @new_mcont@, we have two important observations in the disentangled AAM
where the continuations were represented by data types:
1) When constructing the @new_ndk@, we observe that it explicitly uses @new_frames@ as one
argument of its top @NDCont@ object. The higher-order version of these continuations follows
the same pattern.
2) In the meta-continuation dispatching function @ndcontinue@, the size of the closures set is
decreasing every time while the other parts of meta-continuation are kept. This means in
the higher-order version, we also need an ability to update the top frame of meta-continuation
and keep the rest as the same.

@@

In this section, based on the refunctionalized AAM, we
further present several other transformations toward a big-step abstract
definitional interpreter:

@@

\textit{Call/Return Matching through Refunctionalization and Caching.}
The refunctionalized AAM we obtained has the perfect match on return-flows even though
we do not have a stack model. The higher-order functions representing continuations
already connect all the executions in order. In fact, the continuations (call stack)
of the analyzed language are blended into the call stack of our defining language
(Scala) through refunctionalization, and hence the call/return flow of the analyzed
language is naturally matched.

In fact, we started with an unbounded-stack AAM which already precisely
matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
semantic artifact; the consequence is that we have no place to explicitly save the context
information in the store or on the stack.
However, refunctionalization not only forces the nondeterministic control flows to be higher-order,
but also drives us to use a different caching algorithm.

@@

\section{Back to Direct-Style} \label{directstyle}

In the previous section, we presented a refunctionalized AAM in extended continuation-passing style,
and then simplified it by utilizing the @nd@ operator.
To obtain a definitional abstract interpreter in direct-style, we have multiple choices of how to proceed:

@@

\Citet{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
for small-step AAM to achieve pushdown analysis. This method is both
simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
Based on the AAM we presented in Section \ref{aam}, the only changes in the code required is not
only keeping track of the entry-point expression of the callee, but also holding the target
environment when allocating a continuation addresses. No other pieces of code would need to be
modified. Notably, this change only causes a constant-factor increase in time complexity of
the analysis if the store is widened.

@@

This paper is greatly inspired by the work of~\citet{darais2017abstracting}.
A first surface-level difference is that our presentation uses A-Normal Form $\lambda$-calculus,
instead of plain $\lambda$-calculus, although our work could be easily adapted to handle plain
$\lambda$-calculus as well. Going through the transformation steps from AAMs to ADIs instead of
only considering the final ADI result provides additional insights. In particular, it reveals that
refunctionalization plays an important role for inheriting the call structure (i.e., the model of
the stack) from the meta-language.
In addition, we demonstrate that the explicit use of monads in ADIs is optional, by translating to
direct-style ADIs with delimited control operators in addition to the monadic version of \citet{darais2017abstracting}.
We use the delimited control operators @shift@ and @reset@ \cite{Danvy:1990:AC:91556.91622} as implemented in Scala
\cite{rompf2009implementing}. The correspondence of delimited continuations and monads is well known
\cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions, filinski1994representing}.
